---
title: "Getting Started with DMRnet"
author: "Szymon Nowakowski"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with DMRnet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# This document

The purpose of this vignette is to introduce readers to `DMRnet` package, 
bring them up to speed by providing a simple
use case example and point interested readers towards more informative material.


# DMRnet package

`DMRnet` is an R package for regression and classification with a 
family of model selection algorithms.
`DMRnet` package supports both continous and categorical predictors. 
The overall number of regressors may exceed the number of observations. 

The selected model consists of a subset of numerical regressors and partitions of levels of factors. 

The available model selection algorithms are the following:

- `DMRnet` is the default model selection algorithm that can be used both for $p<n$ and $p \geq n$ scenarios.
- `DMR` is a model selection algorithm that can be used both for $p<n$ scenario only.
- `GLAMER` is a model selection algorithm that can be used both for $p<n$ and $p \geq n$ scenarios. In comparison to `DMRnet`, the step of reordering variables based on their t-statistics is skipped. This is the first partition selection algorithm in literature for which a partition selection consistency has been proven.
- `SOSnet` is the default model selection algorithm that is used both for $p<n$ and $p \geq n$ scenarios for continous-only regressors (with no categorical variables and, consequently, no partition selection). 

Algorithm-wise, the package user has the following choices:

- The user can select `DMRnet` algorithm by calling `DMRnet()` or `cv.DMRnet()` functions.
- The user can select `GLAMER` algorithm by calling `DMRnet()` or `cv.DMRnet()` functions and passing `algorithm="glamer"` parameter.
- The user can select `DMR` algorithm by calling `DMR()` or `cv.DMR()` functions.
- The `SOSnet` algorithm is automatically chosen if `DMRnet()` or `cv.DMRnet()` functions were called with an input consisting of continous-only regressors.

As this vignette is introductory only, and the choice of an algorithm is a somewhat more advanced topic, from now on it is assumed that the user works with `DMRnet` algorithm only. 



```{r setup}
#library(DMRnet)
```
